---
title: "FE515 2022A Assignment 2"
author: "Lorenzo Ausiello"
date: "10/17/2023"
output:
  pdf_document:
    df_print: kable
    fig_caption: yes
    latex_engine: xelatex
---
```{r setup, include=T}
library(knitr)
knitr::opts_chunk$set(echo = T, out.width="80%",out.height="80%", warning = F)
```


# Question 1: 

## 1.1 

Find the attached JPM.csv file. Use as.Date() function to change the first column to Date object.

```{r}
setwd("C:/Users/loaus/OneDrive - stevens.edu/STEVENS/Intro to R/Assignment/Assignment_2")
JPM <- read.csv("FE515_hw2_JPM.csv")

# Changing first column to Date object
JPM$X <- as.Date(JPM$X)
head(JPM)
tail(JPM)
```

## 1.2

Plot the adjusted close price against the date object (i.e. date object as x-axis and close price as y-axis) in red line (require no points). Set the title as JPM, the label for x-axis as Date and the label for y-axis as Adjusted Close Price.

```{r}
plot(JPM$X, JPM$JPM.Adjusted, type = "l", main = "JPM",
xlim = c(as.Date("1970-01-01"), as.Date("1980-08-16")),ylim = c(0,200),
xlab = "Date", ylab = "Adjusted Close Price", col = "red")
```

## 1.3 

Create a scatter plot of close price against open price (i.e. open prices as x-axis, and close prices as y-axis). Set the x label as ”Open Price” and y label as ”Close Price”.

```{r}
plot(JPM$JPM.Open, JPM$JPM.Close, type = "p", main = "Scatter plot: Open Prices vs Close Prices",
xlim = c(0,200),ylim = c(0,200),
xlab = "Open Price", ylab = "Close Price", col = "red")
```

## 1.4 

Use cut() function to divide adjusted close price into 4 intervals. Generate a barplot for the frequencies of these intervals.
.

```{r}
cut <- table(cut(JPM$JPM.Adjusted,4))
cut
barplot(cut, xlab = "Adjusted Close Price", ylab = "Frequency", main = "Barplot Adj.Close Price intervals")
```

## 1.5

Generate a boxplot of volume against the 4 intervals of adjusted close prices.

```{r}
Intervals <- cut(JPM$JPM.Adjusted,breaks = 4)
tail(Intervals)
boxplot(JPM$JPM.Volume ~ Intervals, xlab = "Adjusted Close Price Intervals", ylab = "Volume", main = "Boxplot of Volume by Price Intervals")
```

## 1.6

Use par() function to create a picture of 4 subplots. Gather the 4 figures from 1.2 - 1.5 into ONE single picture. Please arrange the 4 subplots into a 2 by 2 frame, i.e. a frame consists of 2 columns and 2 rows.

```{r}
{par(mfrow = c(2, 2))
  plot(JPM$X, JPM$JPM.Adjusted, type = "l", main = "Adj. Close Price Path",
  xlim = c(as.Date("1970-01-01"), as.Date("1980-08-16")),ylim = c(0,200),
  xlab = "Date", ylab = "Adjusted Close Price", col = "red")
  plot(JPM$JPM.Open, JPM$JPM.Close, type = "p", main = "Scatter plot: Open Prices vs Close Prices",
  xlim = c(0,200),ylim = c(0,200),
  xlab = "Open Price", ylab = "AClose Price", col = "yellow")
  barplot(cut, xlab = "Adjusted Close Price", ylab = "Frequency", main = "Barplot Adj.Close Price intervals")
  boxplot(JPM$JPM.Volume ~ Intervals, xlab = "Adjusted Close Price Intervals", ylab = "Volume",
          main = "Boxplot of Volume by Price Intervals")}
```


# Question 2
Estimate the volume of the unit sphere (which is just 4π/3) by simulation. Suppose we draw N points from three dimensional uniform distribution.

```{r}
N <- 10000
x <- runif(N)
y <- runif(N)
z <- runif(N)

num_inner_points <- 0
for (i in 1:N) {
  if(x[i]^2 + y[i]^2 + z[i]^2 <=1 & x[i]>= 0 & y[i]>= 0 & z[i]>= 0){
    num_inner_points <- num_inner_points + 1
  }
}
#or
num_inner_points <- sum(x^2 + y^2 + z^2 <= 1)

volume_sphere <- num_inner_points/N
volum_unit_spere <- 8 * volume_sphere
volum_unit_spere
```

# Question 3

## 3.1
Implement a Linear Congruential Generator (LCG)
which generates pseudo-random number from uniform distribution using m = 244944, a = 1597, b = 51749.

```{r}
seed <- 1
LCG <- function(n){
m <- 244944
a <- 1597
b <- 51749
x <- rep(NA, n)
x[1] <- (a * seed + b) %% m
for(i in 1:(n-1)){
x[i + 1] <- (a * x[i] + b) %% m
}
seed <<- x[n]
return(x/m)
}

seed<-1
LCG(5)
seed<-1
LCG(5)
seed<-100
LCG(5)
```

## 3.2
Use the LCG in the previous problem, generate 10000 random numbers from chi-square distribution with 10 degrees of freedom (i.e. df = 10), and assign to a variable.

```{r}
seed <- as.numeric(Sys.time())
rnds <- qchisq(LCG(10000),df = 10)
head(rnds)
```

## 3.3
Visualize the resulting sample from 3.2 using a histogram with 40 bins.
```{r}
hist(rnds, nclass = 40)
```
