---
title: "Assignment 4"
author: "Lorenzo Ausiello"
date: "2023-12-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
library(quantmod)
library(quadprog)
```


**QUESTION 1**

## 1.1

Write this problem into the form on page 9 to find f, ui and ci. Then use function constrOptim() to find the optimal portfolio weights x and the optimal value, you can choose starting point as (2, -2, 0).

```{r cars1}

mu <- c(0.0427, 0.0015, 0.0285)
sigma <- matrix(c(0.01, 0.002, 0.001, 0.002, 0.011, 0.003, 0.001, 0.003, 0.02),nrow=3, ncol=3)

f <- function(x) 0.5*t(x)%*%sigma%*%x

ui <- matrix(c(0.0427, 0.0015, 0.0285, -1, -1, -1), 2, 3, byrow = TRUE)
ci <- c(0.05, -1)

constrOptim(c(2,-2,0), f, grad = NULL, ui = ui, ci = ci)$par
constrOptim(c(2,-2,0), f, grad = NULL, ui = ui, ci = ci)$value
```

## 1.2

Write this problem into the form on page 18 to find D, d, A and b. Then use function solve.QP() to find the optimal portfolio weights x and the optimal value.

```{r cars2}

D <- matrix(c(0.01, 0.002, 0.001, 0.002, 0.011, 0.003, 0.001, 0.003, 0.02),nrow=3, ncol=3)
d <- c(0,0,0)
A <- matrix(c(0.0427, 0.0015, 0.0285, -1, -1, -1), 3, 2)
b <- c(0.05, -1)

solve.QP(D, d, A, b)$solution
solve.QP(D, d, A, b)$value
```

**QUESTION 2**

## 2.1

Use linear interpolation to find r(out).

```{r cars3}

t <- c(0.25, 0.5, 1, 2, 3, 5, 7, 10)
r <- c(0.0009, 0.0011, 0.0016, 0.002, 0.0024, 0.0036, 0.0053, 0.0064)
tout <- c(0.75, 1.5, 4, 6, 8)
rout <- approx(t, r, tout, method = "linear")
print(rout)

plot(t, r, type = "o", col = "blue", pch = 16, xlab = "Time", ylab = "Rate", main = "Linear Interpolation")
points(tout, rout$y, col = "red", pch = 16)

```

## 2.2

Use spline interpolation to find rout.

```{r cars4}
rout.s <- spline(t, r, xout = tout, method = "natural")
rout.s$y

plot(t, r, type = "o", col = "blue", pch = 16, xlab = "Time", ylab = "Rate", main = "Spline Interpolation")
points(tout, rout.s$y, col = "green", pch = 16)
```

**QUESTION 3**

Calculate the value from above equation using numerical integration with S0 = 100, K = 100, T = 1, r = 0.05, sigma = 0.2, compare to the call option price from Black-Scholes formula with the same parameters. 

## Numerical integration

```{r cars5}
S <- 100
K <- 100
sigma <- 0.2
r <- 0.05
tau <- 1

# Retrieving the Call Price 
call.option <- function(x){ (S*exp((r-0.5*sigma^2)*tau+sigma*sqrt(tau)*x)-K)*dnorm(x)}
d2 <- (log(S/K)+(r-0.5*sigma^2)*tau)/(sigma*sqrt(tau))
integrate(call.option, -d2, Inf)$value * exp(-r*tau)
```

## Black-Scholes formula

```{r cars6}

bs.call <- function(S, K, tau, sigma, r){
  d2 <- (log(S/K) + (r - 0.5 * sigma^2) * tau) / (sigma * sqrt(tau))
  d1 <- d2 + sigma * sqrt(tau)
  option_price <- S * pnorm(d1) - K * exp(-r * tau)  * pnorm(d2)
  return(option_price)
}

call_price <- bs.call(S, K, tau, sigma, r)
print(call_price)
```